---
title: "The APM Webapp"
author: "Gerrit Versteeg"
date: "30 July 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction of the APM
The APM webapp enables you to analyze the daily air polution patterns at three measuring stations (exploited by an organization called "RIVM"), located in the Hague (the Netherlands). The three locations are Amsterdamse Veerkade, Bleriotlaan en Rebecquestraat. The data for this research is retreived from <https://www.luchtmeetnet.nl/download>.

The user of the Webapp can select a specific polutant and weekday. The Webapp will then calculate the averaged polution levels in ug/m3 over all measurements taken during the period 2015-2016 for that given polutant and weekday. The resulting plot shows the daily polution pattern, using hourly levels and a loess smoother.

All data and coding is available at [Github](https://github.com/GVersteeg/DDP_Wk4_PGA).

## Data Processing
The data for this research is retrieved in the form of a csv-file with measurements for each of the stations. These csv-files are merged into one dataframe.

```{r loading,ECHO = FALSE, warning=FALSE}
library("tidyr", warn.conflicts=FALSE)     ## load tidyr silently
library("dplyr", warn.conflicts=FALSE)     ## load dplyr silently
library("ggplot2", warn.conflicts=FALSE)   ## load ggplot2 silently
library("lubridate", warn.conflicts=FALSE) ## load lubridate silently
library("hms", warn.conflicts=FALSE)       ## load hms silently

setwd("./PolutionMonitor")
DF_AVK <- tbl_df(read.csv2("./data/Export_AVK.csv"))
DF_BL <- tbl_df(read.csv2("./data/Export_BL.csv"))
DF_RS <- tbl_df(read.csv2("./data/Export_RS.csv"))
DF_RIVM <- bind_rows(DF_AVK, DF_BL, DF_RS) ## concatenate station data
```

```{r Peek1,ECHO = TRUE, comment=""}
str(DF_RIVM)                            ## take a peek
```

## Processing the data
The data is pre-processed in such a fashion that the resulting dataframe can easily be used by the Webapp for calculating the averaged polution levels during a certain weekday.
```{r Preprocessing,ECHO=FALSE}
DF_RIVM$locatie <- sub("Den Haag-","", 
                DF_RIVM$locatie)               ## Skip 'Den Haag-'
DF_prep <-                                     ## Create DF_prep
   DF_RIVM %>%                                 ## using DF-RIVM to filter
   rename(ug_m3 = waarde) %>%                  ## rename "waarde"
   mutate(polutant = as.factor(component)) %>% ## rename & set to factor
   mutate(station = as.factor(locatie)) %>%    ## set to format: factor
   mutate(datetime = as.POSIXct(strptime(tijdstip,
                "%Y-%m-%d %H:%M:%S"))) %>%     ## add datetime column
   mutate(date = as.Date(strptime(tijdstip, 
                "%Y-%m-%d"))) %>%              ## add date column
   mutate(weekday = format(date, "%A")) %>%    ## add weekday column
   mutate(time = as.hms(datetime)) %>%         ## add time column
   select(date, time, weekday, station,
               polutant, ug_m3)                ## select relevant cols
```

```{r Peek2, ECHO=TRUE, comment=""}
DF_prep
```

## Example plot (NO2, average Monday)

```{r day-plot, ECHO=FALSE, warning=FALSE, message=FALSE}
in_pol <- "NO2"
in_weekday <- "Monday"
DF_day <-                               ## create DF_month
        DF_prep %>%                     ## using DF-prep to
        filter(weekday == in_weekday, polutant == in_pol) %>%
        group_by(time, station) %>%
        summarize(avg = mean(ug_m3, 
                na.rm = TRUE)) %>%      ## calc. avg waarde
        arrange(time)                   ## sort ascending

h <- ggplot(DF_day, 
        aes(x=time, y=avg, colour=station))  ## setup graphic object
h + geom_point() +
      geom_smooth(span=0.2,se=FALSE) +       ## plot a trendline
      scale_x_time(name="Daily hour") +      ## label X-ticks
      ylab("Polution level (ug/m3)") +       ## label Y-axis
      ggtitle("Daily polution levels The Hague")
```

