---
title: "Tryout. Daily polution patterns"
author: "Gerrit Versteeg"
date: "July 29, 2017"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis
In the report we research the daily air polution patterns at three RIVM measuring stations located in the Hague (Amsterdamse Veerkade, Bleriotlaat en Rebecquestraat).

As to the daily polution pattern for each averaged weekday a clear maximum is visible for the morning traffic rush, but the peek for the evening traffic rush is not detectable. The reason for this is yet unclear.   


## Data Processing
### Loading the data
The data for this research is derived from: https://www.luchtmeetnet.nl/download
by selecting each of the four stations in The Hague (the Netherlands).
The resulting csv-files have been stored in the subdirectory 'data' of the working directory.

Export_AVK.csv (312 kB) - Amsterdamse Veerkade
Export_BL.csv  (353 kB) - Bleriotlaan
Export_RS.csv  (468 kB) - Rebecquestraat

```{r loading,ECHO = TRUE}
library("tidyr", warn.conflicts=FALSE)          ## load tidyr silently
library("dplyr", warn.conflicts=FALSE)          ## load dplyr silently
library("ggplot2", warn.conflicts=FALSE)        ## load ggplot2 silently
library("lubridate", warn.conflicts=FALSE)      ## load lubridate silently
library("hms", warn.conflicts=FALSE)      ## load lubridate silently

setwd("./PolutionMonitor")
dateDownloaded <- date()                        ## register date download
DF_AVK <- tbl_df(read.csv2("./data/Export_AVK.csv"))
DF_BL <- tbl_df(read.csv2("./data/Export_BL.csv"))
DF_RS <- tbl_df(read.csv2("./data/Export_RS.csv"))
DF_RIVM <- bind_rows(DF_AVK, DF_BL, DF_RS)      ## concatenate station data
DF_RIVM                                         ## take a peek
```

The 'coercion'-warnings tell us that there are differences in the levels between the three stations. The files contain:

* Tijdstip      : Indication of date-time (hour) of measurement (e.g. 03:00:00 indicates the averaged measurements averaged over 02:00:01 until 03:00:00)
* Locatie       : Location: name of the measurement station
* Component     : Name of polutant
* Waarde        : Value: observed polutant levels in ug/m3 (hourly average)
* LKI           : Air Quality Index; ranging from 1 to 11 (1 = low air polution)

For more information: https://www.luchtmeetnet.nl/uitleg#


### Preprocessing the data
Not all stations measure polutants at the same moments in time and not all stations measure the same polutants.

```{r Preprocessing,ECHO=TRUE}
# gmm <- intersect(levels(DF_AVK$tijdstip),           ## vul gmm met de gemeen-
#               intersect(levels(DF_BL$tijdstip),     ## schappelijke meetmomenten
#                 levels(DF_RS$tijdstip)))            ## in alle drie meetstations
# 
# gcp <- intersect(levels(DF_AVK$component),          ## vul gcp met de gemeen-
#               intersect(levels(DF_BL$component),    ## schappelijke indicatoren
#                 levels(DF_RS$component)))           ## in alle drie meetstations
DF_RIVM$locatie <- sub("Den Haag-","", 
                DF_RIVM$locatie)            ## Skip 'Den Haag-' part
DF_prep <-                                          ## Create DF_prep
        DF_RIVM %>%                                 ## using DF-RIVM to filter
#        filter(tijdstip %in% gmm) %>%              ## only common dates
#        filter(indicator %in% gcp) %>%             ## only common indicators
        rename(ug_m3 = waarde) %>%                  ## rename "waarde"
        mutate(polutant = as.factor(component)) %>% ## rename & set to factor
        mutate(station = as.factor(locatie)) %>%    ## set to format: factor
        mutate(datetime = as.POSIXct(strptime(tijdstip,
                "%Y-%m-%d %H:%M:%S"))) %>%          ## add datetime column
        mutate(date = as.Date(strptime(tijdstip, 
                "%Y-%m-%d"))) %>%                   ## add date column
        mutate(weekday = format(date, "%A")) %>%     ## add weekday column
        mutate(time = as.hms(datetime)) %>%     ## add time column
        select(date, time, weekday, station,
               polutant, ug_m3)                     ## select relevant cols
```

### Plotting the indicator patterns during the day

```{r day-plot, ECHO=TRUE}
in_pol <- "NO2"
in_weekday <- "Monday"
DF_day <-                                                ## create DF_month
        DF_prep %>%                                      ## using DF-prep to
        filter(weekday == in_weekday, polutant == in_pol) %>%
        group_by(time, station) %>%
        summarize(avg = mean(ug_m3, na.rm = TRUE)) %>%   ## calc. avg waarde
        arrange(time)                                    ## sort ascending

h <- ggplot(DF_day, 
        aes(x=time, y=avg, colour=station))              ## setup graphic object
h + geom_point() +
      geom_smooth(span=0.2,se=FALSE) +                   ## plot a trendline
      scale_x_time(name="Daily hour") +                  ## label X-ticks
      ylab("Polution level (ug/m3)") +                   ## label Y-axis
      ggtitle("Daily polution levels The Hague")         ## title plot
```

